---
title: "Creating Figures"
author: "Geoffrey Millard"
date: "7/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```


```{r setup2, results= FALSE, echo=F, message=F}
# getwd()

# dir()

# setwd("~/ICMGP2022")
```


```{r setup3, results= FALSE, echo=F, message=F}
setwd("~/Projects/Workshops/ICMGP2022_URaPT")
urls  <- readRDS("data/MW_urls.Rdata")
str(urls) #vector of strings, but Rdata can be in any R object class
```


```{r setup4, results= FALSE, echo=F, message=F}
csvs <- urls[c(6,15,26)] #3 are csv
urls <- urls[-c(6,15,26)] #25 are tabbed

DAT1 <- lapply(urls, read.delim, stringsAsFactors = F) #these are the tab-delimited files
DAT2 <- lapply(csvs, read.delim, stringsAsFactors = F, sep = ",") #these are comma-separated

table(unlist(lapply(DAT1, names)))

library(dplyr) #use dplyr for the %>% pipe operator

lapply(DAT1, names) %>% unlist() %>% table() 
lapply(DAT2, names) %>% unlist %>% table 

head(DAT2[[3]])
DAT2[[3]] %>% head

names(DAT2[[3]]) <- names(DAT2[[2]])
lapply(DAT2, names) %>% unlist %>% table #now all of them are the same

library(data.table)
DAT1 <- rbindlist(DAT1)
DAT2 <- rbindlist(DAT2, fill = T)
names(DAT2) == names(DAT1) 

names(DAT2)[[1]] <- names(DAT1)[[1]]

DAT <- rbind(DAT1, DAT2, fill = T)
str(DAT) #77786 observations of fish and mussel tissue
summary(DAT) #not very useful, because all the factors are in character class
```


## Let's get started!
Data: We are going to use the DAT data file from ln 122 in 'Data Scrape in R.Rmd'
package: ggplot2 which is part of the tidyverse package

The DAT data frame contains a lot of data in a long format.  I have found ggplot typically works best with long format data, however many base-R functions (eg. statistics, base plots) work better with wide format data


### Boxplot

```{r}

 
DAT[DAT$Parameter=='Mercury',] %>% group_by(General_Location, Matrix) %>% summarize(n=n(), mean=mean(Result), sd=sd(Result)) %>% kable() %>% kable_styling(bootstrap_options = c('striped', 'hover'), full_width = F)


DAT[DAT$Parameter=='Mercury',] %>% ggplot(aes(x= Matrix, y=Result))+
  geom_boxplot()+
  theme_bw()

names(DAT)
unique(DAT$Matrix)
unique(DAT$Parameter)
```

### Scatterplot

```{r}
Hg_wide <- DAT[DAT$Matrix=='Mussel'] %>% pivot_wider(names_from = Parameter, values_from = Result)

Hg_wide %>% ggplot(aes(y=Mercury, x=Lead, color=Scientific_Name))+
  geom_point()+
  stat_smooth(method = 'glm')+
  xlim(0, 30)+
  ylim(0, 0.60)+
  scale_color_manual(values = viridisLite::viridis(n=3, end=0.8, direction = -1))+
  theme_bw()
```

### Timeseries

```{r}
Hg_wide %>% ggplot(aes(y=Mercury, x=Fiscal_Year, group=Fiscal_Year, color = Scientific_Name))+
  geom_point()+
  geom_hline(yintercept=0.3, color='red')+
  theme_bw()
```

### Timeseries in boxplot form

```{r}
Hg_wide %>% ggplot(aes(y=Mercury, x=Fiscal_Year, group=Fiscal_Year))+
  geom_boxplot()+
  facet_wrap(~Scientific_Name, nrow = 3)+
  theme_bw()
```

### Make your own Theme!!??!??

This can be really helpful when you are formatting figures for publication, presentations, etc.  I modified this theme from {Benjamin Louis}[https://benjaminlouis-stat.fr/en/blog/2020-05-21-astuces-ggplot-rmarkdown/]

```{r ggplot theme, echo=F, warning=F}
theme_ben <- function(base_size = 14) {
  theme_bw(base_size = base_size) %+replace%
    theme(
      # L'ensemble de la figure
      plot.title = element_text(size = rel(1), face = "bold", margin = margin(0,0,5,0), hjust = 0),
      # Zone où se situe le graphique
      # panel.grid.minor = element_blank(),
      # panel.border = element_blank(),
      # Les axes
      axis.title = element_text(size = rel(0.85), face = "bold"),
      axis.text = element_text(size = rel(0.70), face = "bold"),
      axis.line = element_line(color = "black"),
      # La légende
      legend.title = element_text(size = rel(0.75), face = "bold"),
      legend.text = element_text(size = rel(0.50), face = "bold"),
      legend.key = element_rect(fill = "transparent", colour = NA),
      legend.key.size = unit(1.5, "lines"),
      legend.background = element_rect(fill = "transparent", colour = NA),
      # Les étiquettes dans le cas d'un facetting
      strip.background = element_rect(fill = "transparent", color = NA),
      strip.placement = 'outside',
      strip.text = element_text(size = rel(0.85), face = "bold", color = "Black", margin = margin(5,0,5,0))
    )
}

theme_set(theme_ben())

addline_format <- function(x,...){
    gsub('\\s','\n',x)
}
```

Check out how it changed this figure:

```{r}
Hg_wide %>% ggplot(aes(y=Mercury, x=Fiscal_Year, group=Fiscal_Year))+
  geom_boxplot()+
  facet_wrap(~Scientific_Name, nrow = 3)
```

##Exercise
In the time we've got left, make some new figures!  Maybe look at lead, or arsenic.  Look at Oysters instead of Mussels.  This is a big dataset and we have hardly even scratched the surface!!
