---
title: "ICMGP 2022 MusselWatch Scrape Data"
author: "Caitlin Eger"
date: "7/18/2022"
output: html_document
---

```{r setup, include=FALSE, results = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```

## Loading and Managing data
Please download data folder from github at: https://github.com/gmillard756/ICMGP2022_Coding_Chirality/tree/main/data

### The working directory

##### Where is your working directory?
This is the address or path that represents the location where R is currently working. You can only have one working directory active at a time
```{r, results= FALSE}
getwd()
```

##### What is in your working directory?

This is what R can "see" in the directory. If you want to access something that's not in the directory, you'll need to give R more explicit instructions (more on this later). 
```{r , results= FALSE}
dir()
```

Many, many errors that occur happen because a function is trying to access an object that doesn't reference the correct directory or path. Make sure you know where you're working!

### Setting the working directory manually:

To set the working directory in the RStudio IDE, navigate to the "Files" tab >> click on the three dots >> navigate to your data folder >> click "Open"
To encode the working directory: (copy the line and paste to top of your .R script file)
```{r , results= FALSE}
setwd("E:/ICMGP2022/ICMGP2022_Coding_Chirality")
```

#### Read data from a group of `.csv` files

#### Reading data from a url online
This tinyurl is a shortcut to `.txt` file from the MusselWatch dataset at Kachemak, Alaska. You can view it in your browser and then read it directly into R.
```{r, results= FALSE}
url <- "https://tinyurl.com/Kachemak" #this is .txt data from the MusselWatch dataset at Kachemak, AK
Kmak <- read.delim(url)
head(Kmak)
str(Kmak)
```

#### Reading data from a .Rdata file'

This `.Rdata` file contains a list of urls from the MusselWatch dataset.
```{r, results= FALSE}
setwd("E:/ICMGP2022/ICMGP2022_Coding_Chirality")
urls  <- readRDS("data/MW_urls.Rdata")
str(urls) #vector of strings, but Rdata can be in any R object class
urls
```

#### Reading data from a bunch of urls online

Reading data from a list of urls is almost the same as reading in data from a bunch of `.csv` files on your local hard drive. However, in the the urls from the MusselWatch dataset, some of the `.txt` files are in `.csv` format and some are in tab-delimited format, so we want to read them in seperately. Let's make one vector containing urls with the csv-formatted data and the other vector with the tab-delimited datasets.
```{r, results= FALSE}
csvs <- urls[c(6,15,26)] #3 are csv
urls <- urls[-c(6,15,26)] #25 are tabbed
```

Now we can download datasets from all the urls using `lapply()` again. (This step might take a minute).
```{r, results= FALSE}
DAT1 <- lapply(urls, read.delim, stringsAsFactors = F) #these are the tab-delimited files
DAT2 <- lapply(csvs, read.delim, stringsAsFactors = F, sep = ",") #these are comma-separated
```
Each list has the right number of items (25 in DAT1 and 3 in DAT2)

##### Check column names to see if they match up

This is a nested function call that looks at all the names. Nested calls run from the inside out.
```{r, results= FALSE}
table(unlist(lapply(DAT1, names)))
```

##### Using multiple function operations in one step with dplyr

We can also use the pipe operator for the same call, which creates code that is easier to read and understand than nested function calls.

```{r, results= FALSE}
library(dplyr) #use dplyr for the %>% pipe operator

lapply(DAT1, names) %>% unlist() %>% table() 
lapply(DAT2, names) %>% unlist %>% table 
```
All the data frames in `DAT1` have the same names, but the third data frame in object `DAT2` has different names. We can look at `DAT2[[3]]` to see that there are some names with `.` instead of `_` in between words. Here are a couple other examples of how the pipe operator works.

```{r, results= FALSE}
head(DAT2[[3]])
DAT2[[3]] %>% head
```

Because the columns are all in the same order, in this case we can just take the names for one of the data frames 
```{r, results= FALSE}
names(DAT2[[3]]) <- names(DAT2[[2]])
lapply(DAT2, names) %>% unlist %>% table #now all of them are the same
```


We can use the data.table package to zip all the names from the list of names into a single data frame. Now we can stack all the dataframes from the MusselWatch urls into one big dataset and check if DAT1 and DAT2 have the same names:
```{r, results= FALSE}
library(data.table)
DAT1 <- rbindlist(DAT1)
DAT2 <- rbindlist(DAT2, fill = T)
names(DAT2) == names(DAT1) 
```

The first column still has different names, `names(DAT2)[[1]]` and `names(DAT1)[[1]]` don't match each other. All names need to be the same if we want to stack them into one big datset
```{r, results= FALSE}
names(DAT2)[[1]] <- names(DAT1)[[1]]
```

Finally we rbind! (row bind two data frames `DAT1` and `DAT2` together)
```{r, results= FALSE}
DAT <- rbind(DAT1, DAT2, fill = T)
str(DAT) #77786 observations of fish and mussel tissue
summary(DAT) #not very useful, because all the factors are in character class
```

Now we've just created a dataset with 77,786 observations of trace metals in fish and mussel tissue from across the coastal United States! Great work!

#### Quick summary using `summary()` and `table()`

```{r, results= FALSE}
DAT$study <- as.factor(DAT$study)
DAT <- DAT %>% mutate_if(is.character, as.factor) #conditional mutate
summary(DAT) #better!
table(DAT$Parameter) 
```

## Subsetting datasets

##### Splitting a dataset by factor
The `split()` function takes a data frame and makes a list of smaller data frames based on a factor. It's kind of like the opposite of rbindlist
```{r, results= FALSE}
all_params <- split(DAT, DAT$Parameter)
#str(all_params) #this is a list of dataframes again, but we split by parameter instead of by site!
Hg <- all_params[["Mercury"]]
str(Hg)
summary(Hg)

#now we have a nice big mercury dataset!
table(Hg$Matrix) %>% sort #mostly mussels and oysters
```

##### Selecting a whole column

```{r, results= FALSE}
names(Hg)
Hg$Matrix
Hg["Matrix"]
Hg[[8]]
Hg[,8]

```

##### Selecting a whole row

```{r, results= FALSE}
Hg[4,]
```

##### Selecting a single observation

```{r, results= FALSE}
Hg$Matrix[4]
Hg$Matrix[[4]]
Hg[4,8]
```

##### Selecting columns with dplyr

```{r, results= FALSE}
Hg %>% dplyr::select(Matrix)
```

##### Selecting using a selector vector

A selector vector is an index of rows that meet a logical criteria in the form `k <- which(logical == T)`. I always use the variable `k` to denote a selector vector, so that I know it can always be overwritten after being used. You can name your selector vector something more informative if you know you're going to be using it throughout the code.
```{r, results= FALSE}
k <- which(Hg$Matrix == "Oyster")
k
Hg[k,]
Hg$Scientific_Name[k]

```

##### Filtering data with dplyr 

We can also filter all of the rows in the `Hg` dataset to select only samples from mussel species, not other specimens.
```{r, results= FALSE}
Mussels <- Hg %>% dplyr::filter(Matrix == "Mussel")
head(Mussels)
```

## Summarizing datasets

Remember earlier when we summarized the Mussel Watch dataset `DAT` to get a sense what was in those 77K observations? We can do the same again with the mercury subset, and do more complex summaries using dplyr.
```{r, results= FALSE}
summary(Hg) #by column, not by row
```

##### Summarizing by category

There are 6 core dplyr functions: 
  - select: selects, renames and orders columns
  - filter: filters rows by a logical criteria
  - group_by: creates groups by factor 
  - arrange: orders rows in ascending or descending order by one or more columns 
  - summarize: performs a calculation groupwise across groups 
  - mutate: performs a calculation to overwrite or create a new column  

We can combine them using the pipe operator `%>%` to create complex dataset summaries. Let's try some:
```{r, results= FALSE}
Hg %>% group_by(Matrix) %>% summarise(avgHg = mean(Result)) %>% arrange(desc(avgHg))
Hg %>% group_by(Matrix) %>% 
  summarise(avgHg = mean(Result),
            stdevHg = sd(Result),
            n = n()) %>% 
  arrange(desc(avgHg))

Hg %>% group_by(Matrix, Scientific_Name) %>% 
  summarise(avgHg = mean(Result),
            stdevHg = sd(Result),
            n = n()) %>% 
  arrange(desc(avgHg))
```


Can you summarize the Selenium dataset too?
