---
title: "Untitled"
author: "Geoffrey Millard"
date: "7/15/2022"
output: html_document
---

## Importing data
### setup Rmarkdown

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
library(kableExtra)
library(tidyverse)
library(dataRetrieval)
```

### R

```{r set up workspace, warning=F, message=FALSE}
# install.packages("dataRetrieval")
# library(tidyverse)

#USGS and EPA dataset retrieval package
# library(dataRetrieval) 
#click on "Packages" tab to see functions available in the dataRetrival package

#these are the USGS sites upstream from Honnedaga Lake, NY that we found in the browser
sites <- c('0134277112', '0134277114')

#detailed location information about the site
# readNWISsite(c('0134277112', '0134277114'))

#whatNWISdata displays the datasets available at each site
# available <- whatNWISdata(siteNumber=c('0134277112', '0134277114')) # This function no longer works for water quality data
# actual <- readNWISpCode(parameterCd = available$parm_cd) #interpret parameter codes
want <- c('00681', '00945', '50287', '50285') #we want: DOC, DSO4, DHg and DMeHg, 0060 is discharge
# data <- readNWISqw(siteNumbers = sites, parameterCd=want) # get the data.  This function is depricated for water quality data.
Q <- readNWISdv(siteNumbers = sites, parameterCd='00060')
# codes <- readNWISpCode(parameterCd = unique(data$parm_cd)) #confirms that we got what we wanted, shows UNITS!

data <- readWQPqw(paste0('USGS-', sites), want)

sites2 <- c('0134277112', '0134277114', '0425078820', '04250804', '433031074541401', '433202074505900')

siteinfo <- whatWQPdata(siteNumber = paste0('USGS-', sites2)) # seems to return GIS information
# siteinfo
# ?whatWQPdata


# view(data)
# head(data)

```

### Python
In Terminal
$ pip install dataretrieval


```{python, python NWIS import}
import matplotlib.pyplot as plt
import pandas as pd
import dataretrieval.nwis as nwis

# get data from NWIS
site = '0134277112', '0134277114'
```

I have not had great luck reading in data from multiple sites in a single step with this method, but it does pull in all the available parameters for a given site in wide format.

```{python, python NWIS import2}
df12 = nwis.get_record( sites=site[0],  start='2011-05-16', end='2019-10-08', service='qwdata')

df14 = nwis.get_record( sites=r.sites[1], start='2011-05-16', end='2019-10-08', service='dv')

df12b = nwis.get_qwdata(datetime_index = 'TRUE', site_no=r.sites[0])
```


### Exercise
Use either Python or R to grab two physical characteristics:  $Temperature, water$ and $UV 254 -- SDWA NPDWR$

```{r, include=F}
# actual[actual$srsname %in% c('Temperature, water', 'UV 254 -- SDWA NPDWR'),]
cd <- c('00010', '50624')
```

```{python, include=F}
trial12 = nwis.get_record(sites=site[0], start='2011-05-16', end='2019-10-08', service='qwdata', parameterCd = r.cd)
trial14 = nwis.get_record(sites=site[1], start='2011-05-16', end='2019-10-08', service='qwdata', parameterCd = r.cd)
```

```{r, include=F}
trial <- readWQPqw(siteNumbers = paste0('USGS-', sites), parameterCd=cd)
```

### Making a map with leaflet
Install necessary packages.  Leaflet is nice because it allows you to use GIS information from online sources without installing them directly onto your computer.  I have built my own maps using this tutorial[https://waterdata.usgs.gov/blog/basemaps/] from Jason Fisher, USGS
```{r, warning=F}
# install.packages("leaflet")
# install.packages('rgdal')
library(leaflet)
library(rgdal)
```

create a leaflet object
```{r}
map <- leaflet()
```

define options and set up a function
```{r}
lopt <- labelOptions(noHide = TRUE,
                    textOnly = TRUE,
                    style = list('color'='white',
                                 'font-size'='12px'))

GetURL <- function(service, host = "basemap.nationalmap.gov") {
  sprintf("https://%s/arcgis/services/%s/MapServer/WmsServer", host, service)
}

grp <- c("USGS Topo", 
         "USGS Imagery Only", 
         "USGS Imagery Topo",
         "USGS Shaded Relief", 
         "Hydrography")

att <- paste0("<a href='https://www.usgs.gov/'>",
              "U.S. Geological Survey</a> | ",
              "<a href='https://www.usgs.gov/laws/policies_notices.html'>",
              "Policies</a>")
```

Add information (tiles) to the leaflet map
```{r}
# map <- leaflet::addWMSTiles(map, GetURL("USGSTopo"),
#                             group = grp[1], attribution = att, layers = "0")
map <- leaflet::addWMSTiles(map, GetURL("USGSImageryOnly"),
                            group = grp[2], attribution = att, layers = "0")
# map <- leaflet::addWMSTiles(map, GetURL("USGSImageryTopo"),
#                             group = grp[3], attribution = att, layers = "0")
# map <- leaflet::addWMSTiles(map, GetURL("USGSShadedReliefOnly"),
#                             group = grp[4], attribution = att, layers = "0")
# map <- leaflet::addWMSTiles(map, GetURL("USGSHydroCached"),
#                             group = grp[5], options = opt, layers = "0")

```

Pipe different map options
```{r}
# map %>% addCircleMarkers(lat = siteinfo$lat, lng = siteinfo$lon, label = siteinfo$ResolvedMonitoringLocationTypeName, labelOptions = lopt) %>% addScaleBar(position = 'bottomleft', scaleBarOptions(imperial = F))
# 
# map %>% addCircleMarkers(lat = siteinfo$lat, lng = siteinfo$lon, labelOptions = lopt) %>% addScaleBar(position = 'bottomleft', scaleBarOptions(imperial = F))

map %>% addAwesomeMarkers(lat = siteinfo$lat, lng = siteinfo$lon, label = siteinfo$ResolvedMonitoringLocationTypeName, labelOptions = lopt) %>% addScaleBar(position = 'bottomleft', scaleBarOptions(imperial = F))
# 
# map %>% addAwesomeMarkers(lat = siteinfo$lat, lng = siteinfo$lon, label = siteinfo$ResolvedMonitoringLocationTypeName, labelOptions = lopt) %>% addScaleBar(position = 'bottomleft', scaleBarOptions(imperial = F)) %>% fitBounds(lat1 = 35, lat2 = 45, lng1 = -122, lng2 = -70)
```


